\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-09 paper/paper.tex
Updates.

The core network operation in replication is to receive a request from the
client at the leader replica and send that exact same request (as a log entry)
to all follower replicas. The leader therefore \emph{multicasts} a received
request packet to follower replicas, a problem well studied in the networking
community when building software packet switches. Software packet switches reach
impressive speeds of millions of packets forwarded per second~\cite{dpdk_perf}, 
a number far in excess of the few hundreds of packets we manage in
Figure~\ref{fig:problem}. The key insight in removing network overheads in
Cyclone is therefore to approximate a software packet switch for the networking
component of Cyclone.

This is however not a straightforward mapping. Software packet switches make a
simple (usually stateless or sometimes flow associated) decision for each
packet. In contrast, the role of replicas in Cyclone is asymmetric. Log entries
flow from the leader replica to followers and acknowledgments flow
back. Further, the assigned role to a replica can change on reconfiguration
after a failure. A replica must also reject packets from old configurations. The
action associated with a received packet therefore is far more complex than
simply forwarding it to a log.  The \emph{challenge} is to do this efficiently
in software.

\subsection{DPDK}
The Data Plane Development Kit~\cite{dpdk} provides low latency userspace access
to an ethernet NIC, permitting the application to directly send and receive raw
ethernet frames via the transmit and receive queues on the NIC. DPDK is often
used by developers of software packet switches and therefore we leverage
DPDK as a library for building Cyclone.

@@

\subsection{Memory Management}
\label{sec:mm}
We direct DPDK to use pages backed by NVM for packet buffers. DPDK uses a
concurrent memory allocator based on reference counting - used by both the NIC
driver and CPU cores. This means that packet allocation and deallocation does
not happen in the same order as their corresponding position in the replicated
RAFT log. To deal with this, we use a level of indirection as shown in
Figure~\ref{fig:nvm_log}. The NVM log is maintained as a circular log of fixed
sized pointers to the actual packets. Adding a level of indirection in the NVM
log allows us to separate the FIFO ordered circular log being manipulated by
RAFT from packet data being managed by the memory allocator of DPDK. Both the
circular log and packet data are in NVM. An advantage of this scheme is that it
makes recovery from NVM easy -- appends to the circular pointer log are atomic
and we can use the pointer log to recover allocator state i.e. what pieces of
NVM are currently in use by the log.

A key advantage of RAFT is that unlike alternatives such as
MultiPaxos~\cite{multipaxos} or Quorum based replication~\cite{quorum}, the
leader is guaranteed to have the most up to date logs. There is therefore no
case where a leader needs to receive log entries to ``fill'' holes in its logs
from follower replicas, simplifying the protocol state machine we need to
implement. A leader can immediately reject any responses that are not more
current than its view (term). In turn, the persistent log operates as a double
ended queue, with entries either being appended (or possibly deleted at
followers) at the end and being deleted from the front (after commit). This is
critical to efficient operation of Cyclone, as it allows a simple top level
structure (a circular buffer of pointers) to represent the log.

\subsection{Data Movement}
\label{sec:dm}
Software packet switches built on DPDK try to touch as little of the packet data
as possible, minimizing movement of data up and down the cache and memory
hierarchy. Log replication looks like multicast that does not require deep
packet inspection. Software packet switches implementing multicast therefore
simply manipulate packet headers to produce new packets to send on the output
ports. We designed our implementation of the RAFT protocol based on the same
principle.

@@

We now turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
\emph{before} it is multicast out to follower replicas. DPDK userspace NIC
drivers operate in DDIO mode~\cite{ddio} where the packet is directly written
into the CPU cache rather than first being DMAed into DRAM and then fetched by
the CPU on demand. Since the NVM is directly attached we need to persist the
packet to it by executing a cacheline flush ({\tt clflush}) instruction for
every cacheline in the packet and the pointer in the pointer buffer to persist
these via the memory bus. This is not too onerous a burden because we can use
the newly introduced {\tt clflush-opt}~\cite{clflush_opt} instruction
specifically intended to efficiently flush to persistent memory without the
overhead of the serialization normally introduced by {\tt clflush}. This allows
us to hit full memory bandwidth on current generation platforms, a quantity in
excess of 200 Gb/s per core which is well above the near term speeds of 
commodity NICs. We execute a single serializing {\tt sfence} 
before the sequence of {\tt clflushes} to make sure any dirty cachelines due to
header manipulation related writes from the CPU are sent to cache.

\subsection{Batching}

@@

@@

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency is excellent due
to the fact that FPGA cuts out the path from CPU caches to the NIC. Authors
report latency as low as 3us for a round trip from leader to follower replica,
compared to the 6us we observe with DPDK. Notably however they reported 7us when
using TCP/IP from the FPGA rather than their own connection oriented network
protocol. On the other hand their peak replication throughput is 4M
replications/sec on 66 Gbps of aggregate network bandwith for small messages,
comparable to the 6M replications/sec mark obtained by Cyclone on an aggregate
40 Gbps of network bandwidth when \emph{also persisting the log on the SSD}. We
believe the general applicability afforded by our software only approach
together with the fact that we provide persistence by design makes it a
compelling alternative to an FPGA based solution, even given the larger latency.

DARE~\cite{dare} implements state machine replication over RDMA, using a custom
replication protocol. DARE does not consider persistence in its design. It is
worthwhile to point out that since DARE allows any replica to become the leader
it requires a log adjustment protocol for the leader to bring its logs up to
date, a situation we avoid by using RAFT as described in
Section~\ref{sec:mm}. Using 40 Gb/s speeds, and 3-way replication DARE reported
being able to replicate 500K requests/sec. This is without persisting the log of
updates. In contrast, Cyclone replicates upwards of 6 million similarly sized
requests per second with 4*10 Gbps ethernet (without the second level
flashlog). DARE reports an unloaded write latency of 15us, which is lower than
the unloaded write latency of Cyclone at 20us, thanks to infiniband. The
conclusions we can draw are similar to the comparison with Mojim: a carefully
implemented log replication system obviates the need for network offload when
doing replication.

@@

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request is currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores. This is along the lines of similar work~\cite{raft_lin}, using a fixed
client space and maintaining information about last seen sequence numbers from
clients in NVM. Distributed transactions on Cyclone shards is however outside
the scope of this paper.

>>>>

The core network operation in replication is to receive a request from the
client at the leader replica and send that exact same request (as a log entry)
to all follower replicas. The leader therefore \emph{multicasts} a received
request packet to follower replicas, a problem well studied in the networking
community when building software packet switches. Software packet switches reach
impressive speeds of millions of packets forwarded per second~\cite{dpdk_perf}, 
a number far in excess of the few hundreds of packets we manage in
Figure~\ref{fig:problem}. The key insight in removing network overheads in
Cyclone is therefore to approximate a software packet switch for the networking
component of Cyclone. We do this by implementing Cyclone's network stack on top
of the Data Plane Development Kit (DPDK)~\cite{dpdk}, that provides low latency
userspace access to an ethernet NIC. We discuss Cyclone's network protocol in
Section~\ref{sec:netprot}. There are two key challenges in implementing a
consensus protocol as a software packet switch.

Software packet switches make a simple (usually stateless or
sometimes flow associated) decision for each packet. In contrast, running a
consensus protocol requires a relatively complex decision process for each
packet. The direction of flow of log entries depends on the role of the
replicas. Further, the assigned role to a replica can change on reconfiguration
after a failure. A replica must also reject packets from old configurations. In
Section~\ref{sec:dm} we cover the techniques we use in Cyclone to ameliorate
these overheads.

\subsection{DPDK}
\label{sec:netprot}
The Data Plane Development Kit~\cite{dpdk} provides low latency userspace access
to an ethernet NIC, permitting the application to directly send and receive raw
ethernet frames via the transmit and receive queues on the NIC. DPDK is often
used by developers of software packet switches and therefore we leverage
DPDK as a library for building Cyclone.

@@

\subsection{Addressing Overheads}
\label{sec:dm}
Software packet switches built on DPDK try to touch as little of the packet data
as possible, minimizing movement of data up and down the cache and memory
hierarchy. Log replication looks like multicast that does not require deep
packet inspection. Software packet switches implementing multicast therefore
simply manipulate packet headers to produce new packets to send on the output
ports. We designed our implementation of the RAFT protocol based on the same
principle.

@@

RAFT requires that the log entry be persisted \emph{before} it is multicast out
to follower replicas. DPDK userspace NIC drivers operate in DDIO
mode~\cite{ddio} where the packet is directly written into the CPU cache rather
than first being DMAed into DRAM and then fetched by the CPU on demand. Since
the NVM is directly attached we need to persist the packet to it by executing a
cacheline flush ({\tt clflush}) instruction for every cacheline in the packet
and the pointer in the pointer buffer to persist these via the memory bus. This
is not too onerous a burden because we can use the newly introduced {\tt
  clflush-opt}~\cite{clflush_opt} instruction specifically intended to
efficiently flush to persistent memory without the overhead of the serialization
normally introduced by {\tt clflush}. This allows us to hit full memory
bandwidth on current generation platforms, a quantity in excess of 200 Gb/s per
core which is well above the near term speeds of commodity NICs. We execute a
single serializing {\tt sfence} before the sequence of {\tt clflushes} to make
sure any dirty cachelines due to header manipulation related writes from the CPU
are sent to cache.

@@

We direct DPDK to use pages backed by NVM for packet buffers. DPDK uses a
concurrent memory allocator based on reference counting - used by both the NIC
driver and CPU cores. This means that packet allocation and deallocation does
not happen in the same order as their corresponding position in the replicated
RAFT log. To deal with this, we use a level of indirection as shown in
Figure~\ref{fig:nvm_log}. The NVM log is maintained as a circular log of fixed
sized pointers to the actual packets. Adding a level of indirection in the NVM
log allows us to separate the FIFO ordered circular log being manipulated by
RAFT from packet data being managed by the memory allocator of DPDK. Both the
circular log and packet data are in NVM. An advantage of this scheme is that it
makes recovery from NVM easy -- appends to the circular pointer log are atomic
and we can use the pointer log to recover allocator state i.e. what pieces of
NVM are currently in use by the log.

Our choice of RAFT as a consensus protocol is driven by the need to efficiently
maintain this log datastructure. Unlike alternatives such as
MultiPaxos~\cite{multipaxos} or Quorum based replication~\cite{quorum}, the
leader is guaranteed to have the most up to date logs. There is therefore no
case where a leader needs to receive log entries to ``fill'' holes in its logs
from follower replicas, simplifying the protocol state machine we need to
implement. A leader can immediately reject any responses that are not at least as
current as its view (term). In turn, the persistent log operates as a double
ended queue, with entries either being appended (or possibly deleted at
followers) at the end and being deleted from the front (after commit). This is
critical to efficient operation of Cyclone, as it allows a simple top level
structure (a circular buffer of pointers) to represent the log.

@@

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency is excellent due
to the fact that FPGA cuts out the path from CPU caches to the NIC. Authors
report latency as low as 3us for a round trip from leader to follower replica,
compared to the 6us we observe with DPDK. Notably however they reported 7us when
using TCP/IP from the FPGA rather than their own connection oriented network
protocol. On the other hand their peak replication throughput is 4M
replications/sec on 66 Gbps of aggregate network bandwith for small messages,
comparable to the 6M replications/sec mark obtained by Cyclone on an aggregate
40 Gbps of network bandwidth when \emph{also persisting the log on the SSD}. We
believe the general applicability afforded by our software only approach
together with the fact that we provide persistence by design makes it a
compelling alternative to an FPGA based solution, even given the larger
latency. Also, these results demonstrate that a general purpose CPU core can be
as efficient as an FPGA in running a consensus protocol using the techniques
described in this paper.

DARE~\cite{dare} implements state machine replication over RDMA, using a custom
replication protocol. DARE does not consider persistence in its design. It is
worthwhile to point out that since DARE allows any replica to become the leader
it requires a log adjustment protocol for the leader to bring its logs up to
date, a situation we avoid by using RAFT as described in
Section~\ref{sec:dm}. Using 40 Gb/s speeds, and 3-way replication DARE reported
being able to replicate 500K requests/sec. This is without persisting the log of
updates. In contrast, Cyclone replicates upwards of 6 million similarly sized
requests per second with 4*10 Gbps ethernet (without the second level
flashlog). DARE reports an unloaded write latency of 15us, which is lower than
the unloaded write latency of Cyclone at 20us, thanks to infiniband. The
conclusions we can draw are similar to the comparison with Mojim: a carefully
implemented log replication system obviates the need for network offload when
doing replication.

@@

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. We do not provide exactly once semantics as we
believe that exactly once semantics are better provided by the key value store
itself by persisting necessary request markers - as done in systems such as
Kafka~\cite{kafka}. A possible direction of future work however is to add
exactly once semantics to Cyclone's client-server protocol itself to serve as a
building block for distributed transactions across key-value shards, each
running Cyclone, in a manner similar to other systems~\cite{raft_lin}.
